### 2025-04-04 products/manuscript/supplement/Supplementary-Material.qmd

---
title: "Supplement to Effect of COVID-19 burden on changes in HPV vaccination coverage among females between 2019 and 2022"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Supplemental Information

## Table of contents

-   Code and file information
-   Country characteristic details

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

# Country characteristic details
The following list provides additional details for each country-level characteristics, including data source - World Health Organization (WHO) or United Nations (UN)

-   HPV vaccination coverage (WHO)
    -   HPV vaccination coverage in 2022
    -   HPV vaccination coverage in 2019
-   HPV vaccination program characteristics (WHO)
    -   WHO region
    -   World Bank country income level
    -   Indicator for whether HPV vaccine is offered through a national recommendation
    -   Year of HPV vaccine program introduction
    -   Primary HPV vaccine delivery strategy (e.g., school-based)
    -   Indicator for gender neutral vaccination (i.e., female-only vaccination or females and males)
    -   Recommended number of doses (1 or 2)
-   United Nations country demographic data (UN)
    -   Total Population, as of 1 July (thousands)
    -   Population Density, as of 1 July (persons per square km)
    -   Median Age, as of 1 July (years)
    -   Population Growth Rate (percentage)
    -   Total Fertility Rate (live births per woman)
    -   Crude Death Rate (deaths per 1,000 population)
    -   Life Expectancy at Birth, both sexes (years)
    -   Net Migration Rate (per 1,000 population)
-   COVID burden through Dec. 31, 2022 (WHO)
    -   Cumulative number of COVID cases per 100,000 persons
    -   Cumulative number of COVID deaths per 100,000 persons

{{< pagebreak >}}

# Additional Analyses

```{r}
#| label: fig-Figure
#| fig-cap: "Scatterplot of difference in HPV vaccination coverage between 2019 and 2022 and cumulative COVID deaths per 100,000"
#| echo: FALSE
knitr::include_graphics(here("results","figures","scatter_death_covdiff.png"))
```

```{r}
#| label: tbl-medianage_lm_model
#| tbl-cap: "Parameter estimates for median age adjusted linear regression"
#| echo: FALSE
lmtableage_betas = readRDS(here("results", "tables", "lmtableage.rds"))
knitr::kable(lmtableage_betas)
```