### 2025-04-25 project-review-template-Lewis-vandenHurk.qmd
Review 2 - S van den Hurk

---
title: Project Review Template 
author: Shaun van den hurk
date: 25 April 2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Effect of COVID-19 burden on changes in HPV vaccination coverage among females between 2019 and 2022

Name of project author(s): Rayleen Lewis

Name of project reviewer: Shaun van den Hurk

# Specific project content evaluation

Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation

How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The background is very well written and there is clear contextual background in literature and practically. It is clear that the author is very well versed in the subject matter and field. The global burden of HPV, the role of vaccination, and how COVID-19 disrupted immunization services is clearly outlined. The rationale is clear as well as providing a secondary objective of policy utility for the approach provided.

### Summary assessment

-   strong contextualization and motivation

## Question description

How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The primary and secondary research questions are clearly stated and tied directly to the data. The hypothesis is testable and appropriate. The questions logically emerge from the background discussion, and the author appropriately defines the expected direction of the association.

### Summary assessment

-   question/hypotheses fully clear

## Data description

How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The author was very clear on all factors and provided very clear instructions and background. The manuscript gives a clear overview of the data sources (WHO, UN), timeframes, and variable definitions. The integration of datasets using a country identifier is explained, and relevant details (e.g., cumulative COVID burden, vaccination coverage) are well defined. 
The manuscript and description in the EDA components are already clear, but the README file in the raw-data file is particularly clear and helpful.

### Summary assessment

-   source and overall structure of data well explained

## Data wrangling and exploratory analysis

How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Very thorough, well documented and easy to follow. The data wrangling steps are described thoroughly, particularly how the outcome was calculated, how missing data were handled, and how variables were recoded. The EDA is extensive and visually supported through histograms, scatterplots, violin plots, and summary tables, with good descriptions and comments to guide this.

### Summary assessment

-   essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis

Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The project uses a strong combination of methods: univariate, multivariable linear regression, LASSO regression, and regression tree modeling. The justification for each approach is sound, and steps are explained clearly. Performance metrics (R², AIC, BIC) are used to compare models, and tuning is well documented. The author also considers overfitting and model diagnostics.

Additional code was blocked out which is not necessarily but could perhaps be included for the purpose fo the model to use the test/train split approach. But I believe that this is strong and great as a product.

### Summary assessment

-   strong and reasonable analysis

## Presentation

How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Excellent, very high quality, ready for publication. Figures and tables are well labeled, relevant, and appropriately integrated into the narrative. The regression tree is especially useful as a visual summary of subgroup findings.
Excellent use of a reproducible workflow. This is a high quality and finished product which is suitable for a high-end journal and upload in my opinion.

### Summary assessment

-   results are very well presented

## Discussion/Conclusions

Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The discussion appropriately interprets findings and acknowledges the potential for regional or methodological differences. Strengths (e.g., regression tree for policy insight) and limitations (e.g., national-level data, coverage estimate uncertainty) are clearly outlined. The conclusion is consistent with the results and provides practical insight.
The authors expertise are clearly evident.

### Summary assessment

-   strong, complete and clear discussion

## Further comments

The project is really impressive and excellent quality, and it is clear that you have a great deal of experience and expertise in this field. The project demonstrates excellent statistical reasoning and documentation.

There are very few suggestions for practical improvement.
Some very small things:
- Review the spelling in the Word document where things can be highlighted as when I viewed that document I noticed a few small spelling errors that are likely missed in Quarto (e.g nuetral on line 169). I only saw a couple and these will be quick to fix.
- Abstract: Small comment - you started by saying COVID-19 and then just said COVID, i am not sure if this was intentional or if you want to update them all to COVID-19.
- I do not know if train/test split and using CV is important for the sake of the course and if you want to add them (perhaps you did and I missed that). I do not think you need it more generally, but perhaps it is valuable to demonstrate approaches covered in the course.

# Overall project content evaluation

Evaluate overall features of the project by filling in the sections below.

## Structure

Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Everything is clearly organised and the README files were well used and outlined the approach and the code. Files and folders were well used and clealy explained. This exemplifies the concepts outlined in the course.
The only small uncertainty that I had was why there are two eda files but this was covered in the README file. This could perhaps also be added as a line at the top of bboth EDA scripts but is not necessary.

### Summary assessment

-   well structured

## Documentation

How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

The methods section in the manuscript is very well documented. Variable definitions, model choices, and tuning procedures are explained in full, with logical rationale. There is good transparency around decisions, especially with model comparisons. The only suggestion could be to add some of this interpretation and outlining of the approach into the statistical analysis file script, although that is not necessary and will add to the length of the document and affect the clean look of the current document.

### Summary assessment

-   fully and well documented

## Reproducibility

Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Excellent use of a reproducible workflow to run the full analysis. The analysis was also fast to run in its entirety. 
The only very small (tiny) intervention that was needed, was to add "library(patchwork)" to your eda_for_manuscript script so that the script would run fully. This wouldn't be an issues when you are not opening the script directly and was included in the statistical-analysis file.

### Summary assessment

-   fully reproducible without issues

## Thoroughness

How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The author explored multiple modeling approaches, assessed diagnostics, and discussed results with nuance. Exploratory analyses were wide-ranging and the rationale for variable inclusion is clear. Overall, a very comprehensive and thoughtful analysis.
As mentioned, it is clear that the author has expertise in this field.

### Summary assessment

-   strong level of thorougness

## Further comments

Fantastic effort! This is a well-designed, well-documented, and well-executed project. I believe that it exceeds any expectations for the course and is of a very high quality that should be published in a scientific journal. It was clear to me that you have a strong background and experience in these approaches and content. The approach is very logical, clear, efficient and thorough. I learned a lot from reviewing this project and seeing the approach and application of concepts.